---
title: "Transformation Codes"
output: html_notebook:
          self_contained: no
---



```{r}
library(portes)
library(parallel)
library(tseries)
library(readr)
library(urca)
install.packages("Hmisc",
                 repos = c("http://rstudio.org/_packages",
                           "http://cran.rstudio.com"))
```


Variables in levels - testing for stationarity

Real GDP
```{r}
plot(tcodes$X1, type = 'l', xlab = 'year',  ylab = '')
```

clearly growing over time - use adf test with a trend and kpss test for trend stationarity

```{r}
test <- ur.df(tcodes$X1, type = "trend", lags = trunc((length(tcodes$X1)-1)^(1/3)))
summary(test)

kpss.test(tcodes$X1, null = "Trend") #default null: series is level stationary, put kpss.test(x, null = "Trend", lshort = TRUE) for trend stationarity
```

-we strongly reject trend stationarity and we cannot reject phi3 null => it has a unit root with drift
- use tcode 5

--------------------------------------------------------------------------------------------------
--------------------------------------------------------------------------------------------------



S&P/Case-Shiller U.S. National Home Price Index


```{r}
plot(tcodes$X20, type = 'l', xlab = 'year',  ylab = '')
```

clearly growing over time - use adf test with a trend and kpss test for trend stationarity

```{r}
test <- ur.df(tcodes$X20, type = "trend", lags = trunc((length(tcodes$X20)-1)^(1/3)))
summary(test)

kpss.test(tcodes$X20, null = "Trend") #default null: series is level stationary, put kpss.test(x, null = "Trend", lshort = TRUE) for trend stationarity
```

-we strongly reject trend stationarity and we cannot reject phi3 null => it has a unit root with drift
- use tcode 5

--------------------------------------------------------------------------------------------------
--------------------------------------------------------------------------------------------------


Capacity utilization


```{r}
plot(tcodes$X33, type = 'l', xlab = 'year',  ylab = '')
```

clearly not growing over time - use adf test with a drift and kpss test for level stationarity

```{r}
test <- ur.df(tcodes$X33, type = "drift", lags = trunc((length(tcodes$X33)-1)^(1/3)))
summary(test)

kpss.test(tcodes$X33, null = "L") #default null: series is level stationary, put kpss.test(x, null = "Trend", lshort = TRUE) for trend stationarity
```

-phi1 is strongly rejected, tau2 is also, so we conclude that it is stationary -> code 1
(but level stationarity is rejected at 5%)

--------------------------------------------------------------------------------------------------
--------------------------------------------------------------------------------------------------

Civilian unemployment rate


```{r}
plot(tcodes$X34, type = 'l', xlab = 'year',  ylab = '')
```

clearly not growing over time - use adf test with a drift and kpss test for level stationarity

```{r}
test <- ur.df(tcodes$X34, type = "drift", lags = trunc((length(tcodes$X34)-1)^(1/3)))
summary(test)

kpss.test(tcodes$X34, null = "L") #default null: series is level stationary, put kpss.test(x, null = "Trend", lshort = TRUE) for trend stationarity
```

- it is non-stationary, we cannot reject the null plus level stationarity strongly rejected

```{r}
ha <- diff(tcodes$X34)
test <- ur.df(ha, type = "drift", lags = trunc((length(ha)-1)^(1/3)))
summary(test)

kpss.test(ha, null = "L") #default null: series is level stationary, put kpss.test(x, null = "Trend", lshort = TRUE) for trend stationarity
```

-still non-stationary in log first differences

--------------------------------------------------------------------------------------------------
--------------------------------------------------------------------------------------------------

Average hourly earnings

```{r}
plot(tcodes$X36, type = 'l', xlab = 'year',  ylab = '')
```

clearly growing over time - use adf test with a trend and kpss test for trend stationarity

```{r}
test <- ur.df(tcodes$X36, type = "trend", lags = trunc((length(tcodes$X36)-1)^(1/3)))
summary(test)

kpss.test(tcodes$X36, null = "T") #default null: series is level stationary, put kpss.test(x, null = "Trend", lshort = TRUE) for trend stationarity
```

-we cannot reject the null of phi3 => unit root with drift

--------------------------------------------------------------------------------------------------
--------------------------------------------------------------------------------------------------

Real disposable personal income


```{r}
plot(tcodes$X32, type = 'l', xlab = 'year',  ylab = '')
```

clearly growing over time - use adf test with a trend and kpss test for trend stationarity

```{r}
test <- ur.df(tcodes$X32, type = "trend", lags = trunc((length(tcodes$X32)-1)^(1/3)))
summary(test)

kpss.test(tcodes$X32, null = "T") #default null: series is level stationary, put kpss.test(x, null = "Trend", lshort = TRUE) for trend stationarity
```
-the same story as before -> code 5


--------------------------------------------------------------------------------------------------
--------------------------------------------------------------------------------------------------

Real personal consumption expenditures

```{r}
plot(tcodes$X28, type = 'l', xlab = 'year',  ylab = '')
```

clearly growing over time - use adf test with a trend and kpss test for trend stationarity

```{r}
test <- ur.df(tcodes$X28, type = "trend", lags = trunc((length(tcodes$X28)-1)^(1/3)))
summary(test)

kpss.test(tcodes$X28, null = "T") #default null: series is level stationary, put kpss.test(x, null = "Trend", lshort = TRUE) for trend stationarity
```
-unit root with drift -> 5

--------------------------------------------------------------------------------------------------
--------------------------------------------------------------------------------------------------

Real Private Residential Fixed Investment

```{r}
plot(tcodes$X2, type = 'l', xlab = 'year',  ylab = '')
```


-we cannot say if it's growing or not (in the long run, it should be growing)
=> conduct t-test for Case 1: as if it is growing

```{r}
test <- ur.df(tcodes$X2, type = "trend", lags = trunc((length(tcodes$X2)-1)^(1/3)))
summary(test)

kpss.test(tcodes$X2, null = "Trend") #default null: series is level stationary, put kpss.test(x, null = "Trend", lshort = TRUE) for trend stationarity

test2 <- ur.df(tcodes$X2, type = "drift", lags = trunc((length(tcodes$X2)-1)^(1/3)))
summary(test2)

kpss.test(tcodes$X2, null = "Level")

```
- by KPSS test we can strongly reject the null of trend stationarity

from ADF test:
-phi3 null can be rejected at 5% but not at 1% - either there is no unit root and no trend, or unit root plus trend, or no unit root and deterministic trend - the 2nd option is not much plausible, the third option is ruled out by KPSS test for trend stationarity and by test3
-trend is not significant - see test3!!
-tau2: o trochu nemuzeme vyloucit unit root na 5%


-from the first test, we have more evidence that there is not a unit root - rejected at 5%

```{r}
ldif <- diff(log(tcodes$X2))

test2 <- ur.df(ldif, type = "drift", lags = trunc((length(ldif)-1)^(1/3)))
summary(test2)

kpss.test(ldif, null = "Level")
```


Detrend it, use the residuals
```{r}
rinv <- ts(tcodes$X2, start = c(1999,1), frequency = 12)
Trend <- seq(1:length(tcodes$X2))
lag1 <- Lag(rinv, shift = 1)
```


```{r}
model <- lm(rinv ~ Trend)
summary(model)
e <- residuals(model)
```

```{r}
kpss.test(e)
t <- ur.df(e, type = "drift", lags = trunc((length(e)-1)^(1/3)))
summary(t)
```






-we cannot reject the null of a unit root
-use transformation 1, still it is not good


--------------------------------------------------------------------------------------------------
--------------------------------------------------------------------------------------------------

CPI
```{r}
plot(tcodes$X18, type = 'l', xlab = 'year',  ylab = '')
```

clearly growing over time - use adf test with a trend and kpss test for trend stationarity

```{r}
test <- ur.df(tcodes$X18, type = "trend", lags = trunc((length(tcodes$X18)-1)^(1/3)))
summary(test)

kpss.test(tcodes$X18, null = "T") #default null: series is level stationary, put kpss.test(x, null = "Trend", lshort = TRUE) for trend stationarity
```
-strong evidence of a unit root with drift

--------------------------------------------------------------------------------------------------
--------------------------------------------------------------------------------------------------
Mortgage debt outstanding
```{r}
plot(tcodes$X16, type = 'l', xlab = 'year',  ylab = '')
```

-we are not sure if it's growing over time or not => conduct the growing case

```{r}
test <- ur.df(tcodes$X16, type = "trend", lags = trunc((length(tcodes$X16)-1)^(1/3)))
summary(test)

kpss.test(tcodes$X16, null = "T") #default null: series is level stationary, put kpss.test(x, null = "Trend", lshort = TRUE) for trend stationarity
```
-cannot reject phi3 or tau3 => unit root with or without drift

trend in the regression above is not significant:
```{r}
test <- ur.df(tcodes$X16, type = "drift", lags = trunc((length(tcodes$X16)-1)^(1/3)))
summary(test)

kpss.test(tcodes$X16, null = "L") #default null: series is level stationary, put kpss.test(x, null = "Trend", lshort = TRUE) for trend stationarity
```


--------------------------------------------------------------------------------------------------
--------------------------------------------------------------------------------------------------

Housing starts

```{r}
plot(tcodes$X35, type = 'l', xlab = 'year',  ylab = '')
```

-not sure if it's growing => include trend to tests

```{r}
test <- ur.df(tcodes$X35, type = "trend", lags = trunc((length(tcodes$X35)-1)^(1/3)))
summary(test)

kpss.test(tcodes$X35, null = "T") #default null: series is level stationary, put kpss.test(x, null = "Trend", lshort = TRUE) for trend stationarity
```
-cannot reject => unit root with or without drift

trend in the regression above is not significant =>
```{r}
test <- ur.df(tcodes$X35, type = "drift", lags = trunc((length(tcodes$X35)-1)^(1/3)))
summary(test)

kpss.test(tcodes$X35, null = "L") #default null: series is level stationary, put kpss.test(x, null = "Trend", lshort = TRUE) for trend stationarity
```

--------------------------------------------------------------------------------------------------
--------------------------------------------------------------------------------------------------
Total reserves
```{r}
plot(tcodes$X40, type = 'l', xlab = 'year',  ylab = '')
```

don't know if it's growing over time

```{r}
test <- ur.df(tcodes$X40, type = "trend", lags = trunc((length(tcodes$X40)-1)^(1/3)))
summary(test)

kpss.test(tcodes$X40, null = "T") #default null: series is level stationary, put kpss.test(x, null = "Trend", lshort = TRUE) for trend stationarity
```
-same as before: unit root with or without drift

--------------------------------------------------------------------------------------------------
--------------------------------------------------------------------------------------------------

M2
```{r}
plot(tcodes$X27, type = 'l', xlab = 'year',  ylab = '')
```

clearly growing over time - use adf test with a trend and kpss test for trend stationarity

```{r}
test <- ur.df(tcodes$X27, type = "trend", lags = trunc((length(tcodes$X27)-1)^(1/3)))
summary(test)

kpss.test(tcodes$X27, null = "T") #default null: series is level stationary, put kpss.test(x, null = "Trend", lshort = TRUE) for trend stationarity
```
-> unit root with drift

--------------------------------------------------------------------------------------------------
--------------------------------------------------------------------------------------------------
10-year treasury
```{r}
plot(tcodes$X26, type = 'l', xlab = 'year',  ylab = '')
```

don't know if it's growing or not, it may have a decreasing trend during our subsample

```{r}
test <- ur.df(tcodes$X26, type = "trend", lags = trunc((length(tcodes$X26)-1)^(1/3)))
summary(test)

kpss.test(tcodes$X26, null = "T") #default null: series is level stationary, put kpss.test(x, null = "Trend", lshort = TRUE) for trend stationarity
```
-unit root, but we won't use any transformation, since it's a rate

```{r}
test <- ur.df(tcodes$X26, type = "drift", lags = trunc((length(tcodes$X26)-1)^(1/3)))
summary(test)

kpss.test(tcodes$X26, null = "L") #default null: series is level stationary, put kpss.test(x, null = "Trend", lshort = TRUE) for trend stationarity
```

--------------------------------------------------------------------------------------------------
--------------------------------------------------------------------------------------------------
Mortgage average
```{r}
plot(tcodes$X51, type = 'l', xlab = 'year',  ylab = '')
```

don't know

```{r}
test <- ur.df(tcodes$X51, type = "trend", lags = trunc((length(tcodes$X51)-1)^(1/3)))
summary(test)

kpss.test(tcodes$X51, null = "T") #default null: series is level stationary, put kpss.test(x, null = "Trend", lshort = TRUE) for trend stationarity
```
-same as before

```{r}
test <- ur.df(tcodes$X51, type = "drift", lags = trunc((length(tcodes$X51)-1)^(1/3)))
summary(test)

kpss.test(tcodes$X51, null = "L") #default null: series is level stationary, put kpss.test(x, null = "Trend", lshort = TRUE) for trend stationarity
```

--------------------------------------------------------------------------------------------------
--------------------------------------------------------------------------------------------------
Oil price
```{r}
plot(tcodes$X37, type = 'l', xlab = 'year',  ylab = '')
```

don't know

```{r}
test <- ur.df(tcodes$X37, type = "trend", lags = trunc((length(tcodes$X37)-1)^(1/3)))
summary(test)

kpss.test(tcodes$X37, null = "T") #default null: series is level stationary, put kpss.test(x, null = "Trend", lshort = TRUE) for trend stationarity
```
->unit root with or without drift

-trend in the regression above is insignificant=>
```{r}
test <- ur.df(tcodes$X37, type = "drift", lags = trunc((length(tcodes$X37)-1)^(1/3)))
summary(test)

kpss.test(tcodes$X37, null = "L") #default null: series is level stationary, put kpss.test(x, null = "Trend", lshort = TRUE) for trend stationarity
```

--------------------------------------------------------------------------------------------------
--------------------------------------------------------------------------------------------------
S&P 500
```{r}
plot(tcodes$X31, type = 'l', xlab = 'year',  ylab = '')
```

don't know

```{r}
test <- ur.df(tcodes$X31, type = "trend", lags = trunc((length(tcodes$X31)-1)^(1/3)))
summary(test)

kpss.test(tcodes$X31, null = "T") #default null: series is level stationary, put kpss.test(x, null = "Trend", lshort = TRUE) for trend stationarity
```
-> unit root

--------------------------------------------------------------------------------------------------
--------------------------------------------------------------------------------------------------
global price index - commodities
```{r}
plot(tcodes$X49, type = 'l', xlab = 'year',  ylab = '')
```

don't know

```{r}
test <- ur.df(tcodes$X49, type = "trend", lags = trunc((length(tcodes$X49)-1)^(1/3)))
summary(test)

kpss.test(tcodes$X49, null = "T") #default null: series is level stationary, put kpss.test(x, null = "Trend", lshort = TRUE) for trend stationarity
```

-> unit root

-trend is insignificant =>
```{r}
test <- ur.df(tcodes$X49, type = "drift", lags = trunc((length(tcodes$X49)-1)^(1/3)))
summary(test)

kpss.test(tcodes$X49, null = "L") #default null: series is level stationary, put kpss.test(x, null = "Trend", lshort = TRUE) for trend stationarity
```
--------------------------------------------------------------------------------------------------
--------------------------------------------------------------------------------------------------
Real Broad Effective Exchange Rate

```{r}
plot(tcodes$X41, type = 'l', xlab = 'year',  ylab = '')
```
not growing

```{r}
test <- ur.df(tcodes$X41, type = "drift", lags = trunc((length(tcodes$X41)-1)^(1/3)))
summary(test)

kpss.test(tcodes$X41, null = "L") #default null: series is level stationary, put kpss.test(x, null = "Trend", lshort = TRUE) for trend stationarity
```
-> unit root

--------------------------------------------------------------------------------------------------
--------------------------------------------------------------------------------------------------

Volatility
```{r}
plot(tcodes$X52, type = 'l', xlab = 'year',  ylab = '')
```

not growing

```{r}
test <- ur.df(tcodes$X52, type = "drift", lags = trunc((length(tcodes$X52)-1)^(1/3)))
summary(test)

kpss.test(tcodes$X52, null = "L") #default null: series is level stationary, put kpss.test(x, null = "Trend", lshort = TRUE) for trend stationarity
```
-still cannot reject -> unit root

--------------------------------------------------------------------------------------------------
--------------------------------------------------------------------------------------------------

Wu&Xia rate
```{r}
plot(tcodes$X24, type = 'l', xlab = 'year',  ylab = '')
```

```{r}
test <- ur.df(tcodes$X24, type = "trend", lags = trunc((length(tcodes$X24)-1)^(1/3)))
summary(test)

kpss.test(tcodes$X24, null = "T") #default null: series is level stationary, put kpss.test(x, null = "Trend", lshort = TRUE) for trend stationarity
```



```{r}
test <- ur.df(tcodes$X24, type = "drift", lags = trunc((length(tcodes$X24)-1)^(1/3)))
summary(test)

kpss.test(tcodes$X24, null = "L") #default null: series is level stationary, put kpss.test(x, null = "Trend", lshort = TRUE) for trend stationarity
```

Industrial production index
```{r}
plot(ipi_L$X1, type = 'l', xlab = 'year',  ylab = '')
```

-we are unsure whether it is growing or not

```{r}
test <- ur.df(ipi_L$X1, type = "trend", lags = trunc((length(ipi_L$X1)-1)^(1/3)))
summary(test)

kpss.test(ipi_L$X1, null = "T") #default null: series is level stationary, put kpss.test(x, null = "Trend", lshort = TRUE) for trend stationarity
```

Detrend it, use the residuals
```{r}
indpro <- ts(ipi_L$X1, start = c(1999,1), frequency = 12)
Trend <- seq(1:length(ipi_L$X1))
#lag1 <- Lag(rinv, shift = 1)
```


```{r}
model <- lm(indpro ~ Trend)
summary(model)
e <- residuals(model)
```

```{r}
test <- ur.df(e, type = "drift", lags = trunc((length(e)-1)^(1/3)))
summary(test)

kpss.test(e, null = "L") #default null: series is level stationary, put kpss.test(x, null = "Trend", lshort = TRUE) for trend stationarity
```


--------------------------------------------------------------------------------------------------
--------------------------------------------------------------------------------------------------
MA real GDP index
```{r}
plot(ma_gdp_L$X1, type = 'l', xlab = 'year',  ylab = '')
```

-clearly growing

```{r}
test <- ur.df(ma_gdp_L$X1, type = "trend", lags = trunc((length(ma_gdp_L$X1)-1)^(1/3)))
summary(test)

kpss.test(ma_gdp_L$X1, null = "T") #default null: series is level stationary, put kpss.test(x, null = "Trend", lshort = TRUE) for trend stationarity
```



--------------------------------------------------------------------------------------------------
--------------------------------------------------------------------------------------------------
FHFA House Price Index

```{r}
plot(fhfa_hpi_L$X1, type = 'l', xlab = 'year',  ylab = '')
```

-clearly growing

```{r}
test <- ur.df(fhfa_hpi_L$X1, type = "trend", lags = trunc((length(fhfa_hpi_L$X1)-1)^(1/3)))
summary(test)

kpss.test(fhfa_hpi_L$X1, null = "T") #default null: series is level stationary, put kpss.test(x, null = "Trend", lshort = TRUE) for trend stationarity

adf.test
```



--------------------------------------------------------------------------------------------------
--------------------------------------------------------------------------------------------------
Krippner's SSR

```{r}
plot(krippner_ssr_L$X1, type = 'l', xlab = 'year',  ylab = '')
```

-don't know

```{r}
test <- ur.df(krippner_ssr_L$X1, type = "trend", lags = trunc((length(krippner_ssr_L$X1)-1)^(1/3)))
summary(test)

kpss.test(krippner_ssr_L$X1, null = "T") #default null: series is level stationary, put kpss.test(x, null = "Trend", lshort = TRUE) for trend stationarity
```

```{r}
test <- ur.df(krippner_ssr_L$X1, type = "drift", lags = trunc((length(krippner_ssr_L$X1)-1)^(1/3)))
summary(test)

kpss.test(krippner_ssr_L$X1, null = "L") #default null: series is level stationary, put kpss.test(x, null = "Trend", lshort = TRUE) for trend stationarity
```

--------------------------------------------------------------------------------------------------
--------------------------------------------------------------------------------------------------